A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-11 Stickleback_Paper.tex
Supp and revisions to results

In short, the difference between an individual's trait value and the optimum largely determine's that individuals fitness. 
Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

@@

Dispersal occurs both locally along the coastline in the marine habitat 
and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
There is no dispersal directly between lakes.
The lake--ocean migration rate is denoted $m$, 
which we refer to as the rate of \emph{gene flow} between habitats.
Each new individual in the marine habitat has freshwater parents with probability $m$; 
to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
and reflected to stay within the population.
New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
\textbf{The paragraph above has a confusing structure, skipping for now}

@@

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant that was not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in the individual that initially colonized the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in the individual that initially colonized the new lake, 
        and not a captured allele.
\end{enumerate}
%\textbf{I may be conf}

@@

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Since each lake contains 200 individuals, we refer to these migration rates as between 0.01 and 100 migrants per lake, per generation.
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ( $\approx 20,000$ generations).
At very high rates of gene flow, local adaptation was almost impossible.
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
Surprisingly, only a few migrant per generation from the lakes to the marine were needed to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes.
despite being deleterious in the intervening marine habitat --
which allowed populations arriving in new lakes to adapt quickly, reassembling collections of alleles responsible for freshwater adaptation in other lakes.

@@

\begin{figure}
	\begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
  		\caption{ 
        		Mean individual trait values in the marine habitat (blue line),
        		the original lakes (yellow lines; average in orange),
        		and the new lakes (light green lines; average in dark green),
        		across the course of two simulations, with migration rates of
        		\textbf{(top)} $m=5 \times 10^{-5}$ and
        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation per individual
        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
        		Optimal trait values in the two habitats are at $\pm 10$.
                Analogous plots for other migration rates
                are shown in Figures \ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2}.
		}
  		\label{fig:phenotype_ts2}
	\end{center}
\end{figure}

@@

Local adaptation occurred at most migration rates,
as shown in Figure~\ref{fig:MeanPhenotype}.
At the highest migration rate (at which half of each population was composed of migrants),
populations had only slightly different average trait values.
At lower migration rates (10 migrants per generation and below),
populations adapted to local conditions:
as shown in Figures \ref{fig:phenotype_ts2},
\ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2},
freshwater and marine populations diverged
until the trait means were close to the optimal values in each habitat. 
The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
as jumps in the mean trait value,
which move the trait by an amount of order 1 every few hundred generations.
Trait variation within each population was small compared to the difference between populations.
Across all parameter values, differences at around 16 commonly polymorphic sites 
(eight that shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.

@@

\begin{comment}
\paragraph{Standing genetic variation} 
Despite a dramatic difference in the amount of allelic sharing between lakes, 
standing genetic variance (Figure \ref{fig:SGV}) 
was around 0.05 in the freshwater habitat across all but the highest migration rates. 
Concurrently, genetic variation in the marine population 
steadily increased with migration to a similar value. 
On the one hand, it is not surprising that lakes, as a group, 
carry more genetic variation than the marine habitat, 
since population subdivision allows different alleles to become common in each. 
However, it is striking that at $m=.005$, 
the marine habitat carries as much genetic variation, 
despite a lack of any substantial migration load.
\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Realized genetic architecture}

@@

%%%%%%%%%%%%%%%%%%%%%
\subsection*{Origin of introduced freshwater adaptive alleles}

@@

%wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
%this could be nonsense

% at low M we should be able to predict something ---
% at high M we should be able to predict something ----

%Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
%of marine stickleback introduced intro a freshwater environment.
%While this range is heavily effected by all other selections of parameter values.

%(1) There's a threshold of introgression for both rapid adaptation and migration load. 
%(1.5) Maybe refer to the literature for migration load.  
\subsection*{thresholds}

@@

%In a species that commonly is subjected to two general types of selective pressure 
%such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
%reaching a point of pre or post zygotic

%(3) discuss what signals researchers could possibly look out for or connect to real data

@@

\section*{Supplementary material}
\begin{comment}
\begin{figure}
	\begin{center}
  		\includegraphics{Final_Plots/True_Power.pdf}
  		\caption{ 
		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
            \textbf{
                TODO: fix this caption; x-axis label ($F_{ST}$).}
		}
  		\label{fig:Power_FP}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
  		\caption{
            \textbf{Mean phenotypes through time},
            at migration rates of
            \textbf{(top)} 10 and
            \textbf{(bottom)} 1 migrants per lake per generation.
		}
  		\label{fig:pheno_by_time1}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
  		\caption{
            \textbf{Mean phenotypes through time},
            at migration rates of
            \textbf{(top)} 0.1 and
            \textbf{(bottom)} 0.01 migrants per lake per generation.
		}
  		\label{fig:pheno_by_time2}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
        \includegraphics[width=1.0\linewidth]{plotlyPlots/StandingGeneticVariation.png}
  		\caption{
            \textbf{Standing genetic variance},
            computed as XXX write if we leave this in XXX.
		}
        \label{fig:SGV}
	\end{center}
\end{figure}
\end{comment}

>>>>

In short, the difference between an individual's trait value and the optimum largely determines that individual's fitness. 
Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

@@

\textbf{The paragraph below has a confusing structure, skipping for now ... }
Dispersal occurs both locally along the coastline in the marine habitat 
and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
There is no dispersal directly between lakes.
The lake--ocean migration rate is denoted $m$, 
which we refer to as the rate of \emph{gene flow} between habitats.
Each new individual in the marine habitat has freshwater parents with probability $m$; 
to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
and reflected to stay within the population.
New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.

@@

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant that was not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in the individual that initially colonized the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in the individual that initially colonized the new lake, 
        and not a captured allele.
\end{enumerate}

@@

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Since each lake contains 200 individuals, we refer to these migration rates as between 0.01 and 100 migrants per lake, per generation.
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
At very high rates of gene flow, local adaptation was constrained by the large influx of maladaptive alleles.
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
Surprisingly, only a few migrants per generation from the lakes to the marine were needed
to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes -- despite being deleterious in the intervening marine habitat.

@@

@@

\begin{figure}
	\begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
  		\caption{ 
        		Mean individual trait values in the marine habitat (blue line),
        		the original lakes (yellow lines; average in orange),
        		and the new lakes (light green lines; average in dark green),
        		across the course of two simulations, with migration rates of
        		\textbf{(top)} $m=5 \times 10^{-5}$ and
        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation per individual
        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
        		Optimal trait values in the two habitats are at $\pm 10$.
                	Analogous plots for other migration rates
                	are shown in Figure \ref{fig:phenotype_time_supp}.
		}
  		\label{fig:phenotype_time}
	\end{center}
\end{figure}

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at most migration rates with the exception of the highest migration at which half of each population was composed of migrants.
At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
As shown in Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp},
freshwater and marine populations diverged until the trait means were close to the optimal values in each habitat. 
The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_time} as jumps in the mean trait value,
which move the trait by an amount of order 1 every few hundred generations.
Trait variation within each population was small compared to the difference between populations.
Across all parameter values, differences at around 16 commonly polymorphic sites 
(eight that shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.

@@

%\begin{comment}
%\paragraph{Standing genetic variation} 
%Despite a dramatic difference in the amount of allelic sharing between lakes, 
%standing genetic variance (Figure \ref{fig:SGV}) 
%was around 0.05 in the freshwater habitat across all but the highest migration rates. 
%Concurrently, genetic variation in the marine population 
%steadily increased with migration to a similar value. 
%On the one hand, it is not surprising that lakes, as a group, 
%carry more genetic variation than the marine habitat, 
%since population subdivision allows different alleles to become common in each. 
%However, it is striking that at $m=.005$, 
%the marine habitat carries as much genetic variation, 
%despite a lack of any substantial migration load.
%\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
%\end{comment}

\subsection*{Realized genetic architecture}

@@

\subsection*{Origin of introduced freshwater adaptive alleles}

@@

\subsection*{thresholds}

@@

@@

\section*{Supplementary material}

\begin{figure}
	\begin{center}
  		\includegraphics{Final_Plots/True_Power.pdf}
  		\caption{ 
		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
            \textbf{
                TODO: fix this caption; x-axis label ($F_{ST}$).}
		}
  		\label{fig:Power_FP}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
  		\includegraphics[width=1.0\linewidth]{Final_Plots/Pheno_Time_supp.pdf}
  		\caption{
		Mean individual trait values in the marine habitat (blue line),
        		the original lakes (yellow lines; average in orange),
        		and the new lakes (light green lines; average in dark green),
        		across the course of two simulations, with migration rates of
        		\textbf{(A)} $m=5 \times 10^{-3}$,
        		\textbf{(B)} $m=5 \times 10^{-2}$ and
		\textbf{(B)} $m=5 \times 10^{-1}$.
        		(i.e., 1, 10, and 100 migrants per lake per generation, respectively).
        		Optimal trait values in the two habitats are at $\pm 10$.
		}
  		\label{fig:phenotype_time_supp}
	\end{center}
\end{figure}
